Introduction to Statistical Learning R Lab 1
========================================================

Here is an example of making a randomly correlated variable:

```{r}
x=rnorm(50)
y=x+rnorm(50, mean=50, sd=.1)
cor(x,y)
```

And here is a visualization of this correlation

```{r fig.width=7, fig.height=6}
plot(x,y)
```

Here is a sequence of 60 numbers between -pi and pi
```{r}
x=seq(-pi,pi,length=50)
```

Now exploring contour

```{r fig.width=7, fig.height=6}
y=x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels=45,add=T)
```

And another contor figure, this time with a new transform of the data

```{r fig.width=7, fig.height=6}
fa=(f-t(f))/2
contour(x,y,f,nlevels=15)
```

We can also do this with an "image" plot to show a heatmap
```{r fig.width=7, fig.height=6}
image(x,y,fa)
```

Or a 3d perspective plot
```{r fig.width=7, fig.height=6}
persp(x,y,fa)
```

Which can be rotated, I think, with theta
```{r fig.width=7, fig.height=6}
persp(x,y,fa, theta=30)
```

And further rotated with phi
```{r fig.width=7, fig.height=6}
persp(x,y,fa, theta=30, phi=20)
```

Another angle at phi=70

```{r fig.width=7, fig.height=6}
persp(x,y,fa, theta=30, phi=70)
```

And once again with phi=40
```{r fig.width=7, fig.height=6}
persp(x,y,fa, theta=30, phi=40)
```


Loading data
--------

Here are different ways to load data into R. interesting use of the `fix()` command which loads the data into a spreadsheet like window. I commented this out because it blocks execution and doesn't put anything into the result, but it looks like a spreadsheet page.

```{r}
Auto=read.table("~/src/IntroToStatisticalLearningR/data/Auto.data",header=T,na.strings="?")
dim(Auto)
Auto=na.omit(Auto)
dim(Auto)

```

You can also attach data so that it is easier to plot and whatnot, turning cylindars into a dataframe makes the output a little nicer, a boxplot actually!

```{r fig.width=7, fig.height=6}
attach(Auto)
cylinders=as.factor(cylinders)
plot(cylinders, mpg)
```

Lets slap some lipstick on this plot!
```{r fig.width=7, fig.height=6}
plot(cylinders, mpg, col="red", varwidth=T, xlab="cylinders ", ylab="MPG")
```

And we can do a histogram on this variable as well. `col=2` is the same as `col="red"` btw.
```{r fig.width=7, fig.height=6}
hist(mpg,col=2,breaks=15)
```

We could also do a pairs plot which is really nice.
```{r fig.width=7, fig.height=6}
pairs(Auto)
```

And we can limit the pairing to specific variables.
```{r fig.width=7, fig.height=6}
pairs(~mpg + displacement + horsepower + weight + acceleration, Auto)
```

Now this is cool, you can plot variables, and then run identify. In rstudio you click on points you are interested in, then hit `ESC` on the keyboard. This will then give you a list of the points, and name the points on the figure! Cant show it here.

```
plot(horsepower, mpg)
identify(horsepower,mpg,name)
```


FIN!

